---
title: "Lisbon Treaty"
output: html_document
---

```{r setup, include=FALSE}

#Read the Dataset.
Lisbon_data <- read.csv("lisbon-treaty-2-data.csv",header = T)
#snapshot of the dataset.
Lisbon_data[1:10,1:8]
#Dimension of the dataset.
dim(Lisbon_data)

```

The Dataset contains 338 columns and 1002 Rows, where rows contain the data of the individual choices and their age,education,sex etc.As my task is restricted on some factors only so i will not use the whole dataset.I will create a subset of my original dataset where i keep those rows only which is needed for calculation.

```{r}

Lisbon_data2 <- Lisbon_data[,c(1:3,5,6,9,11,128:137,338)]
head(Lisbon_data2)

```

Next i rename the columns of the dataset to better understanding of what they represent so that i can interpret my solution based on that 

```{r}
colnames(Lisbon_data2) <- c("Registered_Voter","oct_vote","oct_type","June_vote",
                            "June_type","EU_Knowledge","Switzerland_Member","Sex",
                            "Marital","Age_type","Class","Region","Area_type",
                            "Working_status","Industry_type",
                            "Working_sector","Education_status","Age")

Lisbon_data2[1:10,1:5]

#Next i group the eu knowledge parameter into 3 section, poor good and average.
Lisbon_data2$EU_Knowledge[Lisbon_data2$EU_Knowledge >7] <- "Good"
Lisbon_data2$EU_Knowledge[Lisbon_data2$EU_Knowledge <= 7 & 
                            Lisbon_data2$EU_Knowledge >= 4] <- "Average"
Lisbon_data2$EU_Knowledge[Lisbon_data2$EU_Knowledge < 4] <- "Poor"

```


```{r}
#except the age attribute all the variable present in my dataset is categorical so i need to transform it into factors.

Lisbon_data2$Registered_Voter <- as.factor(Lisbon_data2$Registered_Voter)
Lisbon_data2$oct_vote <- as.factor(Lisbon_data2$oct_vote)
Lisbon_data2$oct_type <- as.factor(Lisbon_data2$oct_type)
Lisbon_data2$June_vote <- as.factor(Lisbon_data2$June_vote)
Lisbon_data2$June_type <- as.factor(Lisbon_data2$June_type)
Lisbon_data2$EU_Knowledge <- as.factor(Lisbon_data2$EU_Knowledge)
Lisbon_data2$Switzerland_Member <- as.factor(Lisbon_data2$Switzerland_Member)
Lisbon_data2$Sex <- as.factor(Lisbon_data2$Sex)
Lisbon_data2$Marital <- as.factor(Lisbon_data2$Marital)
Lisbon_data2$Age_type <- as.factor(Lisbon_data2$Age_type)
Lisbon_data2$Class <- as.factor(Lisbon_data2$Class)
Lisbon_data2$Region <- as.factor(Lisbon_data2$Region)
Lisbon_data2$Area_type <- as.factor(Lisbon_data2$Area_type)
Lisbon_data2$Working_status <- as.factor(Lisbon_data2$Working_status)
Lisbon_data2$Industry_type <- as.factor(Lisbon_data2$Industry_type)
Lisbon_data2$Working_sector <- as.factor(Lisbon_data2$Working_sector)
Lisbon_data2$Education_status <- as.factor(Lisbon_data2$Education_status)

#In my dataset the categories coded with number, it's good for efficency of the model but now my task is to explain and understand the behaviour of each category for that i have to relabel it into their original names.

levels(Lisbon_data2$Registered_Voter) <- c("Yes","No","Unknown")

levels(Lisbon_data2$oct_vote) <- c("Vote","Didn't Vote","Unknown")

levels(Lisbon_data2$oct_type) <- c("Favour","Against","Didn't Vote")

levels(Lisbon_data2$June_vote) <- c("Vote","Didn't Vote","Unknown")

levels(Lisbon_data2$June_type) <- c("Favour","Against","Didn't Vote")

levels(Lisbon_data2$Switzerland_Member) <- c("True","False","Don't Know")

levels(Lisbon_data2$Sex) <- c("Male","Female")

levels(Lisbon_data2$Marital) <- c("Married","Single","Other")

levels(Lisbon_data2$Age_type) <- c("18-21","22-24","25-34","35-49",
                                              "50-64","65+")

levels(Lisbon_data2$Class) <- c("Administrative","Supervisory","Skilled Manual",
                                "Unskilled manual","Farmers(>50 acre land)",
                                "Farmers(<50 acre land)")

levels(Lisbon_data2$Region) <- c("Dublin","Leinster","Munster","Ulster")

levels(Lisbon_data2$Area_type) <- c("Urban","Rural")

levels(Lisbon_data2$Working_status) <- c("Housewife","Self_Employed","Employee",
                                         "Unemployed(Job Searching)",
                                         "Unemployed(Not searching jobs)",
                                         "Student","Retired")

levels(Lisbon_data2$Industry_type) <- c("Unknown","Construction","Computer/IT",
                                        "Finance","Agriculture","Food Production",
                                        "Government","Leisure","Manufacturing",
                                        "Media","Profesionals","Retail",
                                        "Other Services","Other")

levels(Lisbon_data2$Working_sector) <- c("Unknown","Public","Private")

levels(Lisbon_data2$Education_status) <- c("Primary","Secondary","Third level",
                                           "School/Colllege")

Lisbon_data2$oct_type[is.na(Lisbon_data2$oct_type)] <- "Didn't Vote"
summary(Lisbon_data2)

```

Q.2

```{r}
library(ggplot2)
library(plotly)

attach(Lisbon_data2)


#Bivariate Plot's

ggplot(Lisbon_data2,mapping = aes(oct_vote,fill = oct_type)) + geom_bar() +
   geom_text(aes(label=..count..),stat="count",position=position_stack(0.5)) + 
     ggtitle("Distribution of Votes based on the parameters that and individual voted or not")

#To create a pie chart of the above graph i have to build a table and then plot it.
oct <- as.data.frame(prop.table(table(oct_type)))
plot_ly(oct, labels = ~oct_type, values = ~Freq, 
         type = 'pie',textposition = 'outside',textinfo = 'label+percent') %>%
          layout(title = "Voting percentage by considering those who didn't vote")

#Now let's build a pie chart by considering those individual's only who voted in the referendum.
oct2 <- Lisbon_data2$oct_type[Lisbon_data2$oct_type != "Didn't Vote"]
oct3 <- as.data.frame(table(oct2))
oct3 <- oct3[1:2,]
oct3
plot_ly(oct3, labels = ~oct2, values = ~Freq, 
         type = 'pie',textposition = 'outside',textinfo = 'label+percent') %>%
          layout(title = "Voting percentage after removing those who didn't vote")

```

Q.3

```{r}

ggplot(Lisbon_data2,mapping = aes(oct_type,Age,fill = oct_type))+geom_boxplot()+
  geom_boxplot(width=0.8) + stat_summary(geom="text", fun.y=quantile,
     aes(label=sprintf("%1.1f", ..y..), color=oct_type),
        position=position_nudge(x=0.5), size=3.5) + 
           ggtitle("Distribution of Age according to the Vote's Given")

table(oct_type,Age_type)

ggplot(Lisbon_data2,mapping = aes(oct_vote,fill = Age_type))+geom_bar()+ 
    facet_grid(~Age_type) +
       geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))+
          theme(axis.text.x = element_text(angle = 90)) + 
             ggtitle("Distribution of individual who voted or not based on age group")

ggplot(Lisbon_data2,mapping = aes(oct_type,fill = Age_type)) + geom_bar()+
   facet_grid(~Age_type) + 
      geom_text(aes(label=..count..),
          stat="count",position=position_stack(0.5)) +
             theme(axis.text.x = element_text(angle = 90)) + 
                ggtitle("distribution of voting output based on the age group")

ggplot(Lisbon_data2,mapping = aes(oct_type,fill = Age_type)) + geom_bar()+
   facet_grid(~Age_type) + 
      geom_text(aes(label=round((..count../sum(..count..))*100,2)),
         stat="count",position=position_stack(0.5)) + 
            theme(axis.text.x = element_text(angle = 90)) + 
               ggtitle("Percentage distribution of voting output based on the age group")

Lisbon_data3 <- Lisbon_data2[Lisbon_data2$oct_type != "Didn't Vote",]
head(Lisbon_data3)

#next consider those individual only who voted in the referendum.
ggplot(Lisbon_data3,mapping = aes(oct_type,Age,fill = oct_type))+geom_boxplot()+
  geom_boxplot(width=0.8) + stat_summary(geom="text", fun.y=quantile,
      aes(label=sprintf("%1.1f", ..y..), color=oct_type),
          position=position_nudge(x=0.5), size=3.5) + 
              ggtitle("Distribution of age of the people who voted in the referendum")


ggplot(Lisbon_data3,mapping = aes(oct_type,fill = Age_type)) + geom_bar()+
    facet_grid(~Age_type) + 
       geom_text(aes(label=..count..),stat="count",position=position_stack(0.5)) +
          theme(axis.text.x = element_text(angle = 90)) + 
              ggtitle("distribution of voting output based on the age group")

ggplot(Lisbon_data3,mapping = aes(oct_type,fill = Age_type)) + geom_bar()+
  facet_grid(~Age_type) + geom_text(aes(label=round((..count../sum(..count..))*100,2)),
      stat="count",position=position_stack(0.5)) + 
         theme(axis.text.x = element_text(angle = 90)) + 
            ggtitle("Percentage distribution of voting output based on the age group")

#Next i will create a table for the same.
table(Lisbon_data3$oct_type,Lisbon_data3$Age_type)
prop.table(table(Lisbon_data3$oct_type,Lisbon_data3$Age_type),1)
prop.table(table(Lisbon_data3$oct_type,Lisbon_data3$Age_type),2)

```

Q.4

```{r}

ggplot(Lisbon_data2,mapping = aes(oct_vote,fill = Class)) + geom_bar() +
   facet_grid(~Class) + geom_text(aes(label=..count..),
       stat="count",position=position_stack(0.5)) +
          theme(axis.text.x = element_text(angle = 90)) +
             ggtitle("Distribution of individual who voted or not based on class")

ggplot(Lisbon_data2,mapping = aes(oct_type , fill = Class)) + geom_bar() +
   facet_grid(~Class) + geom_text(aes(label=..count..)
      ,stat="count",position=position_stack(0.5))+
          theme(axis.text.x = element_text(angle = 90)) + 
             ggtitle("distribution of voting output based on the economic class")

ggplot(Lisbon_data2,mapping = aes(oct_type , fill = Class)) + geom_bar() +
    facet_grid(~Class) + geom_text(aes(label=round((..count../sum(..count..))*100,2)),
        stat="count",position=position_stack(0.5)) +
           theme(axis.text.x = element_text(angle = 90)) + 
              ggtitle("Percentage distribution of voting output based on the economic class")

table(oct_type,Class)
prop.table(table(oct_type,Class))

#consider those only who voted in the referendum.
ggplot(Lisbon_data3,mapping = aes(oct_type , fill = Class)) + geom_bar() +
      facet_grid(~Class) + geom_text(aes(label=..count..)
          ,stat="count",position=position_stack(0.5))+
               theme(axis.text.x = element_text(angle = 90)) + 
                    ggtitle("distribution of voting output based on the economic class")

ggplot(Lisbon_data3,mapping = aes(oct_type , fill = Class)) + geom_bar() +
  facet_grid(~Class) + geom_text(aes(label=round((..count../sum(..count..))*100,2)),
        stat="count",position=position_stack(0.5))+
            theme(axis.text.x = element_text(angle = 90)) + 
                 ggtitle("distribution of voting output based on the economic class")

table(Lisbon_data3$oct_type,Lisbon_data3$Class)
prop.table(table(Lisbon_data3$oct_type,Lisbon_data3$Class))


```

Q.5

```{r}

ggplot(Lisbon_data2,mapping = aes(oct_vote,fill = Education_status))+geom_bar()+
    geom_text(aes(label=..count..),stat="count",position=position_stack(0.5)) + 
        facet_grid(~Education_status)+
           theme(axis.text.x = element_text(angle = 90)) + 
              ggtitle("Distribution of individual who voted or not based on Education status")

ggplot(Lisbon_data2,mapping = aes(oct_type,fill = Education_status))+geom_bar()+
    geom_text(aes(label=..count..),stat="count",position=position_stack(0.5)) +  
       facet_grid(~Education_status)+
          theme(axis.text.x = element_text(angle = 90)) + 
              ggtitle("distribution of voting output based on Education status")

ggplot(Lisbon_data2,mapping = aes(oct_type,fill = Education_status))+geom_bar()+
     geom_text(aes(label=round((..count../sum(..count..))*100,2)),
         stat="count",position=position_stack(0.5)) + facet_grid(~Education_status)+
             theme(axis.text.x = element_text(angle = 90)) + 
                ggtitle("Percentage distribution of voting output based on Education status")

attach(Lisbon_data3)
table(oct_type,Education_status)
prop.table(table(oct_type,Education_status))
prop.table(table(oct_type,Education_status),1)
prop.table(table(oct_type,Education_status),2)


ggplot(Lisbon_data3,mapping = aes(oct_type,fill = Education_status))+geom_bar()+
   geom_text(aes(label=..count..),stat="count",position=position_stack(0.5)) +
      facet_grid(~Education_status)+
         theme(axis.text.x = element_text(angle = 90)) + 
            ggtitle("distribution of voting output based on Education status")

ggplot(Lisbon_data3,mapping = aes(oct_type,fill = Education_status))+geom_bar()+
   geom_text(aes(label=round((..count../sum(..count..))*100,2)),
       stat="count",position=position_stack(0.5)) + facet_grid(~Education_status)+
           theme(axis.text.x = element_text(angle = 90)) + 
               ggtitle("Percentage distribution of voting output based on Education status")

```

Q.6

```{r}
ggplot(Lisbon_data2,mapping = aes(oct_vote,fill = EU_Knowledge))+geom_bar()+
   geom_text(aes(label=..count..),stat="count",position=position_stack(0.5)) + 
       facet_grid(~EU_Knowledge)+
           theme(axis.text.x = element_text(angle = 90)) +
             ggtitle("Distribution of individual who voted or not based EU knowledge")


ggplot(Lisbon_data2,mapping = aes(oct_type,fill = EU_Knowledge))+geom_bar()+
   geom_text(aes(label=..count..),stat="count",position=position_stack(0.5)) + 
        facet_grid(~EU_Knowledge) +
            theme(axis.text.x = element_text(angle = 90)) + 
                ggtitle("distribution of voting output based on EU knowledge of individual")
  

ggplot(Lisbon_data2 , mapping = aes(oct_type,fill = EU_Knowledge))+geom_bar()+
   geom_text(aes(label=round((..count../sum(..count..))*100,2)),
      stat="count",position=position_stack(0.5)) + facet_grid(~EU_Knowledge)+
        theme(axis.text.x = element_text(angle = 90)) + 
            ggtitle("Percentage distribution of voting output based on EU knowledge of individual")
attach(Lisbon_data3)
table(oct_type,EU_Knowledge)
prop.table(table(oct_type,EU_Knowledge))
prop.table(table(oct_type,EU_Knowledge),1)
prop.table(table(oct_type,EU_Knowledge),2)

ggplot(Lisbon_data3,mapping = aes(oct_type,fill = EU_Knowledge))+geom_bar()+
    geom_text(aes(label=..count..),stat="count",position=position_stack(0.5)) +
       facet_grid(~EU_Knowledge)+
          theme(axis.text.x = element_text(angle = 90)) + 
             ggtitle("distribution of voting output based on EU Knowledge of individual")

ggplot(Lisbon_data3,mapping = aes(oct_type,fill = EU_Knowledge))+geom_bar()+
   geom_text(aes(label=round((..count../sum(..count..))*100,2)),
      stat="count",position=position_stack(0.5)) + facet_grid(~EU_Knowledge)+
         theme(axis.text.x = element_text(angle = 90)) + 
            ggtitle("Percentage distribution of voting output based on EU Knowledge of individual")

```
Q.7

```{r}

#To analyze is the Eu knowledge of individual have related to the questionare that switzerland is a member of Eu i will use chi square test.  
table(EU_Knowledge,Switzerland_Member)
# H0 = Eu knowledge of individual is independent of switzerland is a member of Eu.
# H1 = Eu knowledge of individual is dependent of switzerland is a member of Eu.
chisq.test(table(EU_Knowledge,Switzerland_Member))
#p value is 2.2e-16 < 0.05 so i have to reject null hypothesis and accept alternate hypothesis.
#That means Eu knowledge of individual is dependent of switzerland is a member of Eu, and to analyze the dependancy i will use some plot's. 

table(EU_Knowledge,Switzerland_Member)
prop.table(table(EU_Knowledge,Switzerland_Member))
prop.table(table(EU_Knowledge,Switzerland_Member),2)
prop.table(table(EU_Knowledge,Switzerland_Member),1)

ggplot(Lisbon_data2,mapping = aes(EU_Knowledge,fill = Switzerland_Member))+geom_bar()+
   geom_text(aes(label=..count..),stat="count",position=position_stack(0.5)) + 
  facet_grid(~Switzerland_Member)+
  theme(axis.text.x = element_text(angle = 90)) + 
  ggtitle("Comparison of EU knowledge of an individual with the output of an questionare that is switzerland is a member of EU or not ")

ggplot(Lisbon_data2,mapping = aes(EU_Knowledge,fill = Switzerland_Member))+geom_bar()+
   geom_text(aes(label=round((..count../sum(..count..))*100,2)),
       stat="count",position=position_stack(0.5)) + facet_grid(~Switzerland_Member)+
           theme(axis.text.x = element_text(angle = 90)) + 
              ggtitle("Comparison in percentage of EU knowledge of an individual with the output of an questionare that is switzerland is a member of EU or not")

ggplot(Lisbon_data2,mapping = aes(oct_vote,fill = Switzerland_Member))+geom_bar()+
   geom_text(aes(label=..count..),stat="count",position=position_stack(0.5)) + 
      facet_grid(~EU_Knowledge)+
         theme(axis.text.x = element_text(angle = 90)) +
            ggtitle("Distribution of individual who voted or not based on both EU knowledge and Switzerland membership questionare output")


ggplot(Lisbon_data2,mapping = aes(oct_type,fill = EU_Knowledge)) + geom_bar() +
   geom_text(aes(label=..count..),stat="count",position=position_stack(0.5)) +  
      facet_grid(~Switzerland_Member) +
         theme(axis.text.x = element_text(angle = 90)) +
            ggtitle("distribution of voting output based on both EU knowledge and Switzerland membership questionare output")


ggplot(Lisbon_data2,mapping = aes(oct_type,fill = EU_Knowledge))+geom_bar()+
    geom_text(aes(label=round((..count../sum(..count..))*100,2)),
        stat="count",position=position_stack(0.5)) + facet_grid(~Switzerland_Member)+
            theme(axis.text.x = element_text(angle = 90)) +
               ggtitle("distribution of voting output based on both EU knowledge and Switzerland membership questionare output")



table(Lisbon_data3$EU_Knowledge,Lisbon_data3$Switzerland_Member)
prop.table(table(Lisbon_data3$EU_Knowledge,Lisbon_data3$Switzerland_Member))
prop.table(table(Lisbon_data3$EU_Knowledge,Lisbon_data3$Switzerland_Member),2)
prop.table(table(Lisbon_data3$EU_Knowledge,Lisbon_data3$Switzerland_Member),1)


ggplot(Lisbon_data3,mapping = aes(oct_type,fill = EU_Knowledge))+geom_bar() +
   geom_text(aes(label=..count..),stat="count",position=position_stack(0.5)) +
       facet_grid(~Switzerland_Member) + theme(axis.text.x = element_text(angle = 90)) + 
         ggtitle("distribution of voting output based on both EU knowledge and Switzerland membership questionare output")


ggplot(Lisbon_data3,mapping = aes(oct_type,fill = EU_Knowledge)) + geom_bar() +
   geom_text(aes(label=round((..count../sum(..count..))*100,2)),stat="count",
       position=position_stack(0.5)) + facet_grid(~Switzerland_Member) +
           theme(axis.text.x = element_text(angle = 90)) + 
              ggtitle("distribution of voting output based on both EU knowledge and Switzerland membership questionare output")

```
